Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2015-10-29 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
minor tweaks from Hilary's feedback

The smooth level vertical (SLEVE) coordinate proposed by \citet{schaer2002} achieves a more regular TF grid in the middle and top of the domain than the BTF coordinate.  The terrain height is split into large-scale and small-scale components, \(h_1\) and \(h_2\), such that \(h = h_1 + h_2\), with each component having a different exponential decay. The transformation is defined as 
\begin{align}
	z &= z^\star + h_1 b_1 + h_2 b_2
\intertext{where the vertical decay functions are given by}
	b_i &= \frac{\sinh \left( \left( H / s_i \right)^n - \left( z^\star / s_i \right)^n \right)}{\sinh \left( H / s_i \right)^n}
\end{align}
with \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.  The exponent \(n\) was introduced by \citet{leuenberger2010} in order to increase cell thickness in the layers nearest the ground, allowing longer timesteps.  \citet{leuenberger2010} found the exponent has an optimal value of \(n = 1.35\).  Choosing \(n = 1\) gives the decay functions used by \citet{schaer2002}.  An example of the SLEVE grid can be seen in figure~\ref{fig:grid}b.

@@

At $t = \SI{10000}{\second}$ the tracer has passed over the mountain.  The horizontal position of the tracer centre can be calculated by integrating along the trajectory to find $t$, the time taken to pass from one side of the mountain to the other:
\begin{align}
	\mathrm{d}t &= \mathrm{d}x / u(x) \\
	t &= \int_0^x \frac{H - h(x)}{u_0 H}\:\mathrm{d}x \\
	t &= \frac{x}{u_0} - \frac{h_0}{16 u_0 H} \left[ 4x + \frac{\sin 2 (\alpha + \beta) x}{\alpha + \beta} \right.+ \nonumber \\
   &\ \left. \frac{\sin 2(\alpha - \beta) x}{\alpha - \beta} + 2 \left( \frac{\sin 2\alpha x}{\alpha} + \frac{\sin 2\beta x}{\beta} \right) \right]
\end{align}
Hence, we find that \(x(t=\SI{10000}{\second}) = \SI{51577.4}{\meter}\).  Because the velocity field is non-divergent, the flow accelerates over mountain ridges and the tracer travels \SI{1577.4}{\meter} further compared to advection in the purely horizontal velocity field.  Tracer height is unchanged downwind of the mountains because advection is along the terrain following coordinate surface.

@@

The cubic upwind-biased advection scheme takes an approach for treating small cut cells that differs from other existing approaches by adjusting weightings to ensure that advection remains upwind-biased near small cells.  The technique is already implemented in OpenFOAM and will be described in greater detail a future publication.  Combined with semi-implicit timestepping and a grid generator that preserves cell length in the direction of the flow, small cells did not impose additional timestep constraints.

>>>>

The smooth level vertical (SLEVE) coordinate proposed by \citet{schaer2002} achieves a more regular TF grid in the middle and top of the domain than the BTF coordinate.  The terrain height is split into large-scale and small-scale components, \(h_1\) and \(h_2\), such that \(h = h_1 + h_2\), with each component having a different exponential decay. The transformation is defined as 
\begin{align}
	z &= z^\star + h_1 b_1 + h_2 b_2
\intertext{where the vertical decay functions are given by}
	b_i &= \frac{\sinh \left( \left( H / s_i \right)^n - \left( z^\star / s_i \right)^n \right)}{\sinh \left( H / s_i \right)^n}
\end{align}
and \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.  The exponent \(n\) was introduced by \citet{leuenberger2010} in order to increase cell thickness in the layers nearest the ground, allowing longer timesteps.  \citet{leuenberger2010} found the exponent has an optimal value of \(n = 1.35\).  Choosing \(n = 1\) gives the decay functions used by \citet{schaer2002}.  An example of the SLEVE grid can be seen in figure~\ref{fig:grid}b.

@@

At $t = \SI{10000}{\second}$ the tracer has passed over the mountain.  The horizontal position of the tracer centre can be calculated by integrating along the trajectory to find $t$, the time taken to pass from one side of the mountain to the other:
\begin{align}
	\mathrm{d}t &= \mathrm{d}x / u(x) \\
	t &= \int_0^x \frac{H - h(x)}{u_0 H}\:\mathrm{d}x \\
	t &= \frac{x}{u_0} - \frac{h_0}{16 u_0 H} \left[ 4x + \frac{\sin 2 (\alpha + \beta) x}{\alpha + \beta} \right.+ \nonumber \\
   &\ \left. \frac{\sin 2(\alpha - \beta) x}{\alpha - \beta} + 2 \left( \frac{\sin 2\alpha x}{\alpha} + \frac{\sin 2\beta x}{\beta} \right) \right]
\end{align}
Hence, we find that \(x(t=\SI{10000}{\second}) = \SI{51577.4}{\meter}\).  Because the velocity field is non-divergent, the flow accelerates over mountain ridges and the tracer travels \SI{1577.4}{\meter} further compared to advection in the purely horizontal velocity field.  Tracer height is unchanged downwind of the mountains because advection is parallel to the surface.

@@

The cubic upwind-biased advection scheme takes an approach for treating small cut cells that differs from other existing approaches by adjusting weightings to ensure that advection remains upwind-biased near small cells.  The implementation of this technique in OpenFOAM is available at \url{https://github.com/hertzsprung/AtmosFOAM/tree/shaw-weller-2015-mwr} and will be described in greater detail a future publication.  Combined with semi-implicit timestepping and a grid generator that preserves cell length in the direction of the flow, small cells did not impose additional timestep constraints.
